Fourmile

### 2022-09-13 fourmile_dataprocess.Rmd
precip and solar rad prism data ready to be shared

## Compare Niwot SNOTEL pirecip to prism pixel

```{r}
niwot_prism <- read.csv(here('data', 'niwot_prism_compare.csv')) %>% 
  filter(!row_number() %in% c(1:10)) %>%
   rename(date = 1,
          prism_mm = 2) %>%
   mutate(date = ymd(date),
          prism_mm = as.numeric(prism_mm))

@@

#filter to look at sept - june
niwot_winter_comp <- niwot_compare %>%
  mutate(month = month(date)) %>%
  filter(!month %in% c(6:8))

summary(niwot_winter_comp)

niwot_prismshift <- niwot_compare %>%
  mutate_at(c("prism_mm"), funs(lead), n = 1) %>%
  mutate(month = month(date)) %>%
  filter(!month %in% c(6:8)) %>%
  mutate(diff_mm = snotel_mm - prism_mm)

summary(niwot_prismshift)

@@

>>>>

## Compare Niwot SNOTEL precip to prism pixel

### Original prism and Niwot SNOTEL precip comaprison

```{r, echo = F}
niwot_prism <- read.csv(here('data', 'niwot_prism_compare.csv')) %>% 
  filter(!row_number() %in% c(1:10)) %>%
   rename(date = 1,
          prism_mm = 2) %>%
   mutate(date = ymd(date),
          prism_mm = as.numeric(prism_mm))

@@

#pbias(niwot_compare$prism_mm, niwot_compare$snotel_mm)

scatter <- ggplot(niwot_compare, aes(x=snotel_mm, y=prism_mm)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Prism vs. Niwot SNOTEL precip')

ggplotly(scatter)

rmse(niwot_compare$prism_mm, niwot_compare$snotel_mm)
summary(niwot_compare$diff_mm)

#filter to look at sept - june
# niwot_winter_comp <- niwot_compare %>%
#   mutate(month = month(date)) %>%
#   filter(!month %in% c(6:8))
# 
# summary(niwot_winter_comp)
```

### Prism shift 1 day and Niwot SNOTEL precip

```{r, echo = F}
niwot_prismshift <- niwot_compare %>%
  mutate_at(c("prism_mm"), funs(lead), n = 1) %>%
  mutate(month = month(date)) %>%
  #filter(!month %in% c(6:8)) %>%
  mutate(diff_mm = snotel_mm - prism_mm) %>%
  na.omit()

#NSE(niwot_prismshift$prism_mm, niwot_prismshift$snotel_mm)
#pbias(niwot_prismshift$prism_mm, niwot_prismshift$snotel_mm)

scatter_shiftday <- ggplot(niwot_prismshift, aes(x=snotel_mm, y=prism_mm)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Prism shift 1 day versus Niwot SNOTEL precip')

ggplotly(scatter_shiftday)

rmse(niwot_prismshift$prism_mm, niwot_prismshift$snotel_mm)
summary(niwot_prismshift$diff_mm)
```

### Prism shift by fraction and Niwot SNOTEL precip

```{r, echo = F}
niwot_prismfrac <- niwot_compare %>%
  mutate(prism_most_mm = prism_mm * 0.974,
         prism_little_mm = prism_mm * 0.026) %>%
  mutate_at(c('prism_most_mm'), list(lead), n = 1) %>%
  mutate(prism_new_mm = prism_most_mm + prism_little_mm,
         diff_prismfrac = snotel_mm - prism_new_mm) %>%
  na.omit()

scatter_frac <- ggplot(niwot_prismfrac, aes(x=snotel_mm, y=prism_new_mm)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Prism shift by 97.4% and Niwot SNOTEL precip')

ggplotly(scatter_frac)

rmse(niwot_prismfrac$prism_new_mm, niwot_prismfrac$snotel_mm)
summary(niwot_prismfrac$diff_prismfrac)
#NSE(niwot_prismfrac$prism_new_mm, niwot_prismfrac$snotel_mm)
#pbias(niwot_prismfrac$prism_new_mm, niwot_prismfrac$snotel_mm)

@@

## Compare prism solar rad to flux solar rad

### Original Prism and Niwot flux tower solar rad

```{r, echo = F}
srad <- solr %>%
  dplyr::select(-1) %>%
  filter(!row_number() %in% c(1:17)) %>%
  replace(.==-9999, NA) %>%
  rename(date = 1,
         prism_w = 2,
         prism_m = 3,
         prism_e = 4,
         gordon_s = 5,
         gordon_n = 6,
         greenlake = 7) %>%
  mutate(date = dmy(date))

niwot_srad <- Niwo %>%
  mutate(date = ymd(date)) %>%
  rename(niwot_flux_mjm2 = 3) %>%
  dplyr::select(-c(watts_per_m2)) %>%
  replace(. == -9999, NA)

## original  
srad_compare <- full_join(srad, niwot_srad, by = 'date') %>%
  dplyr::select(c(date, prism_w, niwot_flux_mjm2)) %>%
  mutate(diff = niwot_flux_mjm2 - prism_w) %>%
  na.omit()

scatter_srad_orig <- ggplot(srad_compare, aes(x=niwot_flux_mjm2, y=prism_w)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Prism versus Niwot flux tower solar rad')

ggplotly(scatter_srad_orig)

rmse(srad_compare$prism_w, srad_compare$niwot_flux_mjm2)
summary(srad_compare$diff)
```

### Prism shift by 1 day versus Niwot flux tower solar rad

```{r, echo = F}
## 1 day shift
srad_shift1day <- srad_compare %>%
  mutate_at(c("prism_w"),list(lead), n = 1) %>%
  na.omit() %>%
  mutate(diff = niwot_flux_mjm2 - prism_w)

scatter_srad_1day <- ggplot(srad_shift1day, aes(x=niwot_flux_mjm2, y=prism_w)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Prism shift 1 day vs. Niwot flux tower solar rad')

ggplotly(scatter_srad_1day)
#scatter_srad_1day

rmse(srad_shift1day$prism_w, srad_shift1day$niwot_flux_mjm2)
summary(srad_shift1day$diff)
#pbias(srad_shift1day$prism_w, srad_shift1day$niwot_flux_mjm2)
```

### Prism shift by fraction versus Niwot flux tower solar rad

```{r, echo = F}
## shift by percentage
srad_fracshift <- srad_compare %>%
    mutate(prism_most = prism_w * 0.3,
         prism_little = prism_w * 0.7) %>%
  mutate_at(c('prism_most'), list(lead), n = 1) %>%
  mutate(prism_new = prism_most + prism_little,
         diff_prismfrac = niwot_flux_mjm2 - prism_new) %>%
  na.omit()

scatter_srad_frac <- ggplot(srad_fracshift, aes(x=niwot_flux_mjm2, y=prism_new)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Prism shift by 30% vs. Niwot flux tower solar rad')

ggplotly(scatter_srad_frac)
#scatter_srad_frac

rmse(srad_fracshift$prism_new, srad_fracshift$niwot_flux_mjm2)
summary(srad_fracshift$diff_prismfrac)
#pbias(srad_fracshift$prism_new, srad_fracshift$niwot_flux_mjm2)
```

### Daymet (original) versus Niwot flux tower solar rad

```{r, echo = F}
## bring in daymet
## Daily total radiation (MJ/m2/day) can be calculated as follows: ((srad (W/m2) * dayl (s/day)) / l,000,000)
niwot_daymet <- daym %>%
  filter(!row_number() %in% c(1:7)) %>%
  rename(year = 1,
         doy = 2,
         srad_wm2 = 3) %>%
  mutate(date = as.Date(doy),
         month = month(date),
         day = day(date),
         date = make_date(year, month, day),
         daymet_mjm2 = (srad_wm2 * 86400) / 1000000) %>%
  dplyr::select(-c(year, doy, month, day))

niwot_daymet_comp <- full_join(niwot_srad, niwot_daymet, by = 'date') %>%
  dplyr::select(c(date, niwot_flux_mjm2, daymet_mjm2)) %>%
  na.omit() %>%
  mutate(diff = niwot_flux_mjm2 - daymet_mjm2)

scatter_srad_daymet <- ggplot(niwot_daymet_comp, aes(x=niwot_flux_mjm2, y=daymet_mjm2)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Daymet vs. Niwot flux tower solar rad')

ggplotly(scatter_srad_daymet)
#scatter_srad_frac

#summary(srad_fracshift)
rmse(niwot_daymet_comp$daymet_mjm2, niwot_daymet_comp$niwot_flux_mjm2)
summary(niwot_daymet_comp$diff)
```
